---
title: " Modern Data Mining, HW 4"
author:
- Group Member Christian Benitez
- Group Member Taurean Butler
- Group Member 3
date: '11:59 pm, 03/19, 2023'
output:
  pdf_document:
    number_sections: yes
    toc: yes
    toc_depth: '4'
  html_document:
    code_folding: show
    highlight: haddock
    number_sections: yes
    theme: lumen
    toc: yes
    toc_depth: 4
    toc_float: yes
  word_document:
    toc: yes
    toc_depth: '4'
urlcolor: blue
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, results = "hide", fig.width=8, fig.height=4)
options(scipen = 0, digits = 3)  # controls base R output
if(!require('pacman')) {install.packages('pacman')}
pacman::p_load(bestglm, glmnet, leaps, car, tidyverse, pROC, caret,kableExtra) # add the packages needed
```

\pagebreak

# Overview

Logistic regression is used for modeling categorical response variables. The simplest scenario is how to identify risk factors of heart disease? In this case the response takes a possible value of `YES` or `NO`. Logit link function is used to connect the probability of one being a heart disease with other potential risk factors such as `blood pressure`, `cholestrol level`, `weight`. Maximum likelihood function is used to estimate unknown parameters. Inference is made based on the properties of MLE. We use AIC to help nailing down a useful final model. Predictions in categorical response case is also termed as `Classification` problems. One immediately application of logistic regression is to provide a simple yet powerful classification boundaries. Various metrics/criteria are proposed to evaluate the quality of a classification rule such as `False Positive`, `FDR` or `Mis-Classification Errors`. 

LASSO with logistic regression is a powerful tool to get dimension reduction. 


## Objectives

- Understand the model
  - logit function
    + interpretation
  - Likelihood function
- Methods
    - Maximum likelihood estimators
        + Z-intervals/tests
        + Chi-squared likelihood ratio tests
- Metrics/criteria 
    - Sensitivity/False Positive
    - True Positive Prediction/FDR
    - Misclassification Error/Weighted MCE
    - Residual deviance
    - Training/Testing errors

- LASSO 

- R functions/Packages
    - `glm()`, `Anova`
    - `pROC`
    - `cv.glmnet`
  
## R Markdown / Knitr tips

You should think of this R Markdown file as generating a polished report, one that you would be happy to show other people (or your boss). There shouldn't be any extraneous output; all graphs and code run should clearly have a reason to be run. That means that any output in the final file should have explanations.

A few tips:

* Keep each chunk to only output one thing! In R, if you're not doing an assignment (with the `<-` operator), it's probably going to print something.
* If you don't want to print the R code you wrote (but want to run it, and want to show the results), use a chunk declaration like this: `{r, echo=F}`. Notice this is set as a global option. 
* If you don't want to show the results of the R code or the original code, use a chunk declaration like: `{r, include=F}`
* If you don't want to show the results, but show the original code, use a chunk declaration like: `{r, results='hide'}`.
* If you don't want to run the R code in a chunk at all use `{r, eval = F}`.
* We show a few examples of these options in the below example code. 
* For more details about these R Markdown options, see the [documentation](http://yihui.name/knitr/options/).
* Delete the instructions and this R Markdown section, since they're not part of your overall report.

## Review

Review the code and concepts covered in

* Module Logistic Regressions/Classification
* Module LASSO in Logistic Regression

## This homework

We have two parts in this homework. Part I is guided portion of work, designed to get familiar with elements of logistic regressions/classification. Part II, we bring you projects. You have options to choose one topic among either Credit Risk via LendingClub or Diabetes and Health Management. Find details in the projects. 



# Part I: Framingham heart disease study 

We will continue to use the Framingham Data (`Framingham.dat`) so that you are already familiar with the data and the variables. All the results are obtained through training data. 

Liz is a patient with the following readings: `AGE=50, GENDER=FEMALE, SBP=110, DBP=80, CHOL=180, FRW=105, CIG=0`. We would be interested to predict Liz's outcome in heart disease. 

To keep our answers consistent, use a subset of the data, and exclude anyone with a missing entry. For your convenience, we've loaded it here together with a brief summary about the data.

```{r data preparation, include=F}
# Notice that we hide the code and the results here
# Using `include=F` in the chunk declaration. 
hd_data <- read.csv("~/R_Projects/class-codes/Spring 2023/STAT5710/Homework/HW4/Framingham.dat")
str(hd_data) 

### Renames, setting the variables with correct natures...
names(hd_data)[1] <- "HD"
hd_data$HD <- as.factor(hd_data$HD)
hd_data$SEX <- as.factor(hd_data$SEX)
str(hd_data)
#tail(hd_data, 1)    # The last row is for prediction
hd_data.new <- hd_data[1407,] # The female whose HD will be predicted.
hd_data <- hd_data[-1407,]  # take out the last row 
hd_data.f <- na.omit(hd_data)
```

We note that this dataset contains 311 people diagnosed with heart disease and 1095 without heart disease.
```{r table heart disease, echo = F, comment = " ", results = T}
# we use echo = F to avoid showing this R code
# notice the usage of comment = " " here in the header
table(hd_data$HD) # HD: 311 of "0" and 1095 "1" 
```

After a quick cleaning up here is a summary about the data:
```{r data summary, comment=" "}
# using the comment="     ", we get rid of the ## in the output.
summary(hd_data.f)
```

Lastly we would like to show five observations randomly chosen. 
```{r, results = T, comment=" "}
row.names(hd_data.f) <- 1:1393
set.seed(471)
indx <- sample(1393, 5)
hd_data.f[indx, ]
# set.seed(471)
# hd_data.f[sample(1393, 5), ]
```

## Identify risk factors

### Understand the likelihood function
Conceptual questions to understand the building blocks of logistic regression. All the codes in this part should be hidden. We will use a small subset to run a logistic regression of `HD` vs. `SBP`. 

i. Take a random subsample of size 5 from `hd_data_f` which only includes `HD` and `SBP`. Also set  `set.seed(471)`. List the five observations neatly below. No code should be shown here.

```{r, echo=F}
set.seed(471)

hd_data.f %>% select(HD,SBP)%>% sample_n(size = 5,replace = F) %>% kbl(caption = "Sample of 5 Observations of hd_data_f") %>%
  kable_classic(full_width = T, html_font = "Cambria")
```


ii. Write down the likelihood function using the five observations above.

$$\begin{split}
\mathcal{L}(\beta_0, \beta_1 \vert {\text Data}) &= {Prob\text {(the outcome of the data)}}\\
&=Prob((HD=0|SBP=128), (HD=0|SBP=220), (HD=0|SBP=174),Prob(HD=1|SBP=128), Prob(HD=1|SBP=260) \\
&=Prob((HD=0|SBP=128)\times (HD=0|SBP=220)\times (HD=0|SBP=174)\times Prob(HD=1|SBP=128)\times Prob(HD=1|SBP=260) ) \\
&= \frac{1}{1+e^{\beta_0 + 128 \beta_1}}\cdot\frac{1}{1+e^{\beta_0 + 220\beta_1}}\cdot\frac{1}{1+e^{\beta_0 + 174\beta_1}}\cdot\frac{e^{\beta_0 + 128 \beta_1}}{1 + e^{\beta_0 + 128 \beta_1}}\cdot\frac{e^{\beta_0 + 260 \beta_1}}{1 + e^{\beta_0 + 260 \beta_1}} 
	\end{split}$$


iii. Find the MLE based on this subset using glm(). Report the estimated logit function of `SBP` and the probability of `HD`=1. Briefly explain how the MLE are obtained based on ii. above.

$$\begin{split} 

logit = -3.65489 + 0.01581*SBP\\
 \\
 
 \\
\hat P(HD = 1 \vert SBP) &= \frac{e^{-3.66+0.0159 \times  SBP}}{1+e^{-3.66+0.0159 \times SBP}} \\

	\end{split}$$

**In problem ii, we created a likelihood function using our five observations above. Maxmimum likelihood estimates are obtained when you try to find the \beta_0, \beta_1 that maximize the log of the function above in ii.**

```{r}
HD.fit.01 <- glm(HD~SBP, data = hd_data.f, family=binomial(logit))
summary(HD.fit.01)
```


iv. Evaluate the probability of Liz having heart disease. 

**The probability that Liz has a heart disease is .112 or 11.2%**

```{r}
fit1.predict <- predict(HD.fit.01, hd_data.new, type="response") 
fit1.predict
```


### Identify important risk factors for `Heart.Disease.`

We focus on understanding the elements of basic inference method in this part. Let us start a fit with just one factor, `SBP`, and call it `fit1`. We then add one variable to this at a time from among the rest of the variables. For example
```{r, results='hide'}
summary(fit1 <- glm(HD~SBP, hd_data.f, family=binomial))

summary(fit1.1 <- glm(HD~SBP + AGE, hd_data.f, family=binomial))

summary(fit1.2 <- glm(HD~SBP + AGE + SEX, hd_data.f, family=binomial))

summary(fit1.3 <- glm(HD~SBP + AGE + SEX + DBP, hd_data.f, family=binomial))

summary(fit1.4 <- glm(HD~SBP + AGE + SEX + DBP + CHOL, hd_data.f, family=binomial))

summary(fit1.5 <- glm(HD~SBP + AGE + SEX + DBP + CHOL + FRW, hd_data.f, family=binomial))

summary(fit1.6 <- glm(HD~SBP + AGE + SEX + DBP + CHOL + FRW + CIG, hd_data.f, family=binomial))

# you will need to finish by adding each other variable 
# fit1.2...
```

i. Which single variable would be the most important to add?  Add it to your model, and call the new fit `fit2`.  

We will pick up the variable either with highest $|z|$ value, or smallest $p$ value. Report the summary of your `fit2` Note: One way to keep your output neat, we will suggest you using `xtable`. And here is the summary report looks like.

**We chose the value _SEX_ since it has the smallest p value**


```{r the most important addition, results='asis', comment="   "}
## How to control the summary(fit2) output to cut some junk?
## We could use packages: xtable or broom. 
## Assume the fit2 is obtained by SBP + AGE
library(xtable)
options(xtable.comment = FALSE)
fit2 <- glm(HD~SBP + SEX, hd_data.f, family=binomial)
xtable(fit2)
```

ii. Is the residual deviance of `fit2` always smaller than that of `fit1`? Why or why not?

**The residual deviance of fit 2 will be smaller than that of fit1 because we are adding additional parameters into our model. When doing this, our model becomes more flexible and can thus better capture any patterns within the data.**
  
iii. Perform both the Wald test and the Likelihood ratio tests (Chi-Squared) to see if the added variable is significant at the .01 level.  What are the p-values from each test? Are they the same? 

__The p value for SEX from the Wald Test is 1.0e-10 *** and the Chi-Squared test is 1.48e-22 ***. Though they are smaller than the .01 level, they are not same__

```{r}
# Wald Test
summary(fit2)
confint.default(fit2) 
Anova(fit2)

# Likelihood Ratio Test

chi.sq <- 1469.3-1373.8
pvalue <- pchisq(chi.sq, 1, lower.tail=FALSE)
pvalue
```



###  Model building

Start with all variables. Our goal is to fit a well-fitting model, that is still small and easy to interpret (parsimonious).

i. Use backward selection method. Only keep variables whose coefficients are significantly different from 0 at .05 level. Kick out the variable with the largest p-value first, and then re-fit the model to see if there are other variables you want to kick out.

```{r}
summary(fit.all <-glm(HD~SBP + AGE + SEX + DBP + CHOL + FRW + CIG, hd_data.f, family=binomial)) # All variables included

summary(fit.6 <-glm(HD~SBP + AGE + SEX + CHOL + FRW + CIG, hd_data.f, family=binomial)) # Remove DBP

summary(fit.5 <-glm(HD~SBP + AGE + SEX + CHOL + CIG, hd_data.f, family=binomial)) # Remove FRW

summary(fit.4 <-glm(HD~SBP + AGE + SEX + CHOL, hd_data.f, family=binomial)) # Remove CIG

```


ii. Use AIC as the criterion for model selection. Find a model with small AIC through exhaustive search. Does exhaustive search  guarantee that the p-values for all the remaining variables are less than .05? Is our final model here the same as the model from backwards elimination? 

**The model with the smallest AIC value is from model fit.6 which containes the variables SBP, AGE, SEXMALE, CHOL, FRW, CIG. This exhaustive search does not guarantee that the p values for all remaining variables are less than .05.  However, this differs from the final model generated by the backwards elimitation model which is fit.4 which contains SBP, AGE, SEX, CHOL.**

```{r}
Xy_design <- model.matrix(HD ~.+0, hd_data.f) 
Xy <- data.frame(Xy_design, hd_data.f$HD)   
fit.all.AIC <- bestglm(Xy, family = binomial, method = "exhaustive", IC="AIC", nvmax = 10)
fit.all.AIC$BestModel
```



iii. Use the model chosen from part ii. as the final model. Write a brief summary to describe important factors relating to Heart Diseases (i.e. the relationships between those variables in the model and heart disease). Give a definition of “important factors”. 

**Based on our final model, we find that SBP, AGE, SEXMALE, CHOL, FRW, CIG are important factors that related to Heart Disease. SBP is Systolic Blood Pressure, AGE is the age of the subject, SEXMALE is their gender, CHOL is the cholesterol level, FRW is the age and gender adjusted for weight, and CIG is a self-report measure of how many cigarretes are smoked each week. Since we used AIC as our model selection criterion, important factors are defined as the predictors that produce a model that fits the data best compared to other models. Our final model suggests that older adults, males individuals with higher SBP CHOL and more cigarette uses increases your chances of heart disease **

iv. What is the probability that Liz will have heart disease, according to our final model?

**According to our final model, the probability that Liz will have heart disease is 3.46%**

```{r}
Liz02.predict <- predict(fit.6, hd_data.new, type="response") 
Liz02.predict
```

##  Classification analysis

### ROC/FDR

i. Display the ROC curve using `fit1`. Explain what ROC reports and how to use the graph. Specify the classifier such that the False Positive rate is less than .1 and the True Positive rate is as high as possible.

The ROC reports the predictive power of a given model. The closer to the right top diagonal we are, the higher sensitivity/true positive rate (ability to accurately identify that someone actually has heart disease) we have and lower false positives are being made. However, the more to the left bottom diagonal we are, the less predictive power our classifier has (i.e. we have more false positives and less true positives). The curve can also be used to identify an acceptable threshold for when to classify someone as having heart disease. However that does come with a cost. By increasing the threshold, we have a chance of identifying someone as having heart disease when they really don't.

When we specify the classifier such that the False Positive rate is less than .1 and the True Positive rate is as high as possible we get the below result of .298

```{r}
fit1.roc <- roc(hd_data.f$HD, fit1$fitted) 
plot(1-fit1.roc$specificities, 
     fit1.roc$sensitivities, col="red", lwd=3, type="l",
     xlab="False Positive", 
     ylab="Sensitivity")
abline(coef = c(0,1))

coords <- coords(fit1.roc, x = "all", input = "threshold", ret = "all")
coords <- coords[coords$fpr <= 0.1,]
classifier <- coords[which.max(coords$tpr),]
classifier[,c(1:3,12:13)]
```


```{r}

fit.final.roc <- roc(fram_data.f$HD, fit.final$fitted)


lines(1-fit.final.roc$specificities, fit.final.roc$sensitivities, col="green", lwd=3)
legend("bottomright",
       c(paste0("fit1 AUC=", round(fit1.roc$auc,2)), 
         paste0("fit2 AUC=", round(fit2.roc$auc, 2)), 
         paste0("fit.final AUC=", round(fit.final.roc$auc, 2))),
       col=c("red", "blue", "green"),
       lty=1)
```


ii. Overlay two ROC curves: one from `fit1`, the other from `fit2`. Does one curve always contain the other curve? Is the AUC of one curve always larger than the AUC of the other one? Why or why not?

No. The curves are not always contained by the other curve, meaning that the AUC of one model is not always larger over the other model. This can be seen by the bottom left corner showing that the curves appear to be overlapping. The AUC represents the area under the curve, which tells us how our model performs across all classification thresholds. When the AUC is larger, it signifies that the model is able to detect more true positives. From these two models it's a little less clear as to which model out performs depending on the threshold set. Overall, fit2 appears to be slightly better. 


```{r}
fit2.roc <- roc(hd_data.f$HD, fit2$fitted)
plot(1-fit1.roc$specificities, 
     fit1.roc$sensitivities, col="red", lwd=3, type="l",
     xlab="False Positive", 
     ylab="Sensitivity")
lines(1-fit2.roc$specificities, fit2.roc$sensitivities, col="blue", lwd=3)
abline(coef = c(0,1))
legend("bottomright",
       c(paste0("fit1 AUC=", round(fit1.roc$auc,2)), 
         paste0("fit2 AUC=", round(fit2.roc$auc, 2))),
       col=c("red", "blue"),
       lty=1)
```


iii.  Estimate the Positive Prediction Values and Negative Prediction Values for `fit1` and `fit2` using .5 as a threshold. Which model is more desirable if we prioritize the Positive Prediction values?

**If we prioritize the positive prediction value, then model 2 performs better because it has a higher ppv of .472 compared to model 1 which has .45**

```{r}
fit1.pred.5 <- ifelse(fit1$fitted > 1/2, "1", "0")
cm1.5 <- table(fit1.pred.5, hd_data.f$HD)
positive.pred.1 <- cm1.5[2,2] / sum(cm1.5[2,])
negative.pred.1 <- cm1.5[1,1] / sum(cm1.5[1,])

Fit1 <- data.frame(positive.pred.1,negative.pred.1)


fit2.pred.5 <- ifelse(fit2$fitted > 1/2, "1", "0")
cm2.5 <- table(fit2.pred.5, hd_data.f$HD)
positive.pred.2 <- cm2.5[2,2] / sum(cm2.5[2,])
negative.pred.2 <- cm2.5[1,1] / sum(cm2.5[1,])

Fit2 <- data.frame(positive.pred.2,negative.pred.2)

cbind(Fit1,Fit2)


```



iv.  For `fit1`: overlay two curves,  but put the threshold over the probability function as the x-axis and positive prediction values and the negative prediction values as the y-axis.  Overlay the same plot for `fit2`. Which model would you choose if the set of positive and negative prediction values are the concerns? If you can find an R package to do so, you may use it directly.

**Based on the results below, we would choose fit2 because the average number of positive and negative prediction values is larger than that of fit 1**

```{r}
coords01 <- coords(fit1.roc, x = "all", input = "threshold", ret = "all") %>% select(threshold,ppv,npv)

plot(coords01$threshold, coords01$ppv,
     col="red", lwd=3, type="l",
     xlab="threshold", 
     ylab="prediction values", xlim=c(.1,.8), ylim=c(0,1))
lines(coords01$threshold, coords01$npv,col="blue", type = "l")
legend("bottomright",
       c(paste0("ppv"), 
         paste0("npv")),
       col=c("red", "blue"),
       lty=1)

coords02 <- coords(fit2.roc, x = "all", input = "threshold", ret = "all") %>% select(threshold,ppv,npv)

plot(coords02$threshold, coords02$ppv,
     col="red", lwd=3, type="l",
     xlab="threshold", 
     ylab="prediction values", xlim=c(.1,.8), ylim=c(0,1))
lines(coords02$threshold, coords02$npv,col="blue", type = "l")
legend("bottomright",
       c(paste0("ppv"), 
         paste0("npv")),
       col=c("red", "blue"),
       lty=1)

mean(coords01$ppv, na.rm =T)
mean(coords01$npv, na.rm =T)
mean(coords02$ppv, na.rm =T)
mean(coords02$npv, na.rm =T)
```

  
### Cost function/ Bayes Rule

Bayes rules with risk ratio $\frac{a_{10}}{a_{01}}=10$ or $\frac{a_{10}}{a_{01}}=1$. Use your final model obtained from Part 1 to build a class of linear classifiers.


i.  Write down the linear boundary for the Bayes classifier if the risk ratio of $a_{10}/a_{01}=10$.

The linear function for the final model generated from part 1 is: 

$$\begin{split} 

logit = -4.57026  + 0.01872*SBP +  0.90342 *SEXMALE\\

	\end{split}$$
	
The Bayes rule for thresholding is:
	
$$\begin{split} 

\hat P(Y=1 \vert x) > \frac{0.1}{(1+0.1)}=0.09 \\

	\end{split}$$
	
Therefore, the linear boundary is 

$$\hat HD =1 ~~~~ \text{if} ~~~~ SBP > SEX * \frac{-0.90342 }{0.01872} +\frac{4.57026  + log(0.09/.91)}{0.01872}$$
$$  =  SEXMALE * -48.3  + 121$$

```{r}
summary(fit2 <- glm(HD~SBP + SEX, hd_data.f, family=binomial))
```


ii. What is your estimated weighted misclassification error for this given risk ratio?

Our estimated misclassificiation error is .745

```{r}
fit2.pred.bayes <- as.factor(ifelse(fit2$fitted > .09, "1", "0"))
MCE.bayes <- (10*sum(fit2.pred.bayes[hd_data.f$HD == "1"] != "1")
              + sum(fit2.pred.bayes[hd_data.f$HD == "0"] != "0"))/length(hd_data.f$HD)
MCE.bayes
```


iii.  How would you classify Liz under this classifier?

Liz has an SPD of 110. Given our classifier she would not have heart disease since the threshold for females is 121. 

iv. Bayes rule gives us the best rule if we can estimate the probability of `HD-1` accurately. In practice we use logistic regression as our working model. How well does the Bayes rule work in practice? We hope to show in this example it works pretty well.

Now, draw two estimated curves where x = threshold, and y = misclassification errors, corresponding to the thresholding rule given in x-axis.

```{r}
threshold_10 <- seq(0,1, by = .01)
MCE_10 <- seq(0,1, by = .01)
n=length(threshold_10)
for(i in 1:n){
  fit2.pred.bayes <- as.factor(ifelse(fit2$fitted > threshold_10[i], "1", "0"))
  MCE.bayes <- (10*sum(fit2.pred.bayes[hd_data.f$HD == "1"] != "1")
               + sum(fit2.pred.bayes[hd_data.f$HD == "0"] != "0"))/length(hd_data.f$HD)
  MCE_10[i] <- MCE.bayes
}
bayes.10 <- data.frame(threshold_10,MCE_10)

threshold_1 <- seq(0,1, by = .01)
MCE_1 <- seq(0,1, by = .01)
n=length(threshold_1)
for(i in 1:n){
  fit2.pred.bayes <- as.factor(ifelse(fit2$fitted > threshold_1[i], "1", "0"))
  MCE.bayes <- (1*sum(fit2.pred.bayes[hd_data.f$HD == "1"] != "1")
               + sum(fit2.pred.bayes[hd_data.f$HD == "0"] != "0"))/length(hd_data.f$HD)
  MCE_1[i] <- MCE.bayes
}
bayes.1 <- data.frame(threshold_1,MCE_1)
```


v. Use weighted misclassification error, and set $a_{10}/a_{01}=10$. How well does the Bayes rule classifier perform? 

The Bayes rule classifier has an MCE lower than the majority thresholds (about 5%) 

```{r}
bayes.10 %>% ggplot(aes(x = threshold_10, y = MCE_10)) +
  geom_line() + geom_hline(yintercept = .745, col  = "red")

sum(bayes.10$MCE_10<.745)/nrow(bayes.10)

```


vi. Use weighted misclassification error, and set $a_{10}/a_{01}=1$. How well does the Bayes rule classifier perform? 

The Bayes rule classifier has an MCE higher than the majority thresholds (about 91%) 

```{r}
bayes.1 %>% ggplot(aes(x = threshold_1, y = MCE_1)) +
  geom_line() + geom_hline(yintercept = .745, col  = "red")

1 - sum(bayes.1$MCE_1>.745)/nrow(bayes.1)
```



# Part II: Project

## Project Option 1 Credit Risk via LendingClub

## Project Opetion 2  Diabetes and Health Management
